---
title: "p8105_hw1_xy2395"
author: "Jack Yan"
date: "9/13/2018"
output: github_document
---
# Problem 1
## Creating a data frame with different variable types
First, try to create a data frame. The code chunk below creates a data frame comprised of:

1.  A random sample of size 10 from a uniform[0, 5] distribution;
2.  A logical vector indicating whether elements of the sample are greater than 2;
3.  A (length-10) character vector;
4.  A (length-10) factor vector.
```{r}
set.seed(1)
random_sample = runif(10, 0, 5)  
logical_vec = random_sample > 2
char_vec = rep(c("a", "b", "c", "d", "e"), 2)
factor_vec = factor(rep(c(0, 1), 5), levels = c(0, 1))

#create a data frame from the 4 vectors above
data_frame = data.frame(random_sample, logical_vec, char_vec, factor_vec)

#check the data frame
data_frame
```

## Calculating the mean
Now try to take the mean of each variable in the data frame. It apparently works to get the mean of the numeric vector. The `mean()` function also works on logical vectors, as it converts `TRUE` and `FALSE` to `1` and `0`, respectively, and then calculate their arithmetic mean. Code and results are shown below.
```{r}
mean(data_frame$random_sample)
mean(data_frame$logical_vec)
```

In the attempts to get the means of the character vector ,and also the factor vector, a warning shows up and "NA" is returned. This is because `mean()` function doesn't support these two types of variable. 
```{r}
mean(data_frame$char_vec)
mean(data_frame$factor_vec)
```

## Coercion
Now coerce objects into type "numeric". Logical values `FALSE` and `TRUE` are turned into `0` and `1`, respectively. Characters are all turned to `NA`. Factor objects are coerced to numbers indicating their different levels. Results are not shown.
```{r, eval = FALSE}
as.numeric(logical_vec)
as.numeric(char_vec)
as.numeric(factor_vec)
```

Next, try to convert the character variable from character to factor to numeric; similarly, convert the factor variable from factor to character to numeric. As a result, the character variable can be converted to factor to numeric, and different characters are designated different numbers. Similarly, the factor variable can also be converted to character to numeric. 
```{r}
as.numeric(as.factor(char_vec))
as.character(factor_vec)
as.numeric(as.character(factor_vec))
```
However, in another case, if the factor vector is labeled with characters, it will first be converted to the label name, and therefore cannot be converted to numerics, so `NA` will be returned. See below: 
```{r}
factor_vec_labeled = factor(rep(c(0, 1), 5), levels = c(0, 1), labels = c("male","female"))
as.character(factor_vec_labeled)
as.numeric(as.character(factor_vec_labeled))
```


# Problem 2
## Creating a data frame
```{r, include = FALSE}
rm(list=ls())
```
Create a data frame comprised of:

1.  x: a random sample of size 1000 from a standard Normal distribution;
2.  y: a random sample of size 1000 from a standard Normal distribution;
3.  A logical vector indicating whether the x + y > 0;
4.  A numeric vector created by coercing the above logical vector;
5.  A factor vector created by coercing the above logical vector.


```{r}
set.seed(1)
x = rnorm(1000) 
y = rnorm(1000)
positive_sum = x + y > 0
numeric_vec = as.numeric(positive_sum)
factor_vec = as.factor(positive_sum)

#combine the 5 vectors above into data frame
data_frame = data.frame(x, y, positive_sum, numeric_vec, factor_vec)
head(data_frame)
```


The size of the dataset is `r nrow(data_frame)` rows * `r ncol(data_frame)` columns. The mean of x is `r mean(x)`, and the median of x is `r median(x)`. The proportion of cases for which the logical vector is `TRUE` is `r  sum(positive_sum)/length(positive_sum)`.

## Graphing
### Part 1
```{r, include = FALSE}
library(tidyverse)
```
Make a scatterplot of `y` vs `x`, and color points using the logical variable `positive_sum`. Export this scatterplot to my project directory using `ggsave`. The plot will be saved to a .pdf file called `scatter_plot.pdf`.
```{r}
ggplot(data_frame, aes(x = x, y = y, color = positive_sum)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
```

### Part 2
Now make a second and third scatterplot that color points using the numeric and factor variables, respectively.
```{r}
ggplot(data_frame, aes(x = x, y = y, color = numeric_vec)) + geom_point()
ggplot(data_frame, aes(x = x, y = y, color = factor_vec)) + geom_point()
```

Based on the two graphs above, if the points are colored using numeric variables, the color scale is shown as a continuous gradient ranging from the minimum to the maximum. If the points are colored by factor variables, the color scale is shown as separate color dots showing different levels of the factor variables. 


