---
title: "p8105_hw1_xy2395"
author: "Jack Yan"
date: "9/13/2018"
output: github_document
---
# Problem 1
## Creating a data frame with different variable types
First, try to create a data frame. The code chunk above creates a data frame comprised of:

1.  A random sample of size 10 from a uniform[0, 5] distribution;
2.  A logical vector indicating whether elements of the sample are greater than 2;
3.  A (length-10) character vector;
4.  A (length-10) factor vector.
```{r}
set.seed(1)
random_sample = runif(10, 0, 5)  
logical_vec = random_sample > 2
char_vec = rep(c("a", "b", "c", "d", "e"), 2)
factor_vec = factor(rep(c(0, 1), 5), levels = c(0, 1))

#create a data frame from the 4 vectors above
data_frame = data.frame(random_sample, logical_vec, char_vec, factor_vec)
```

## Calculating the mean
Now try to take the mean of each variable in the data frame. It apparently works to get the mean of the numeric vector. The `mean()` function also works on logical vactors, because it converts `TRUE` and `FALSE` to `1` and `0`, respectively, and then calculate their arithmetic mean.
```{r}
mean(data_frame$random_sample)
mean(data_frame$logical_vec)
```

In the attempts to get the means of the character vector, and the factor vector, a warning shows up and "NA" is returned. This is because `mean()` function doesn't support these two types of data. 
```{r}
mean(data_frame$char_vec)
mean(data_frame$factor_vec)
```

## Coercion
Now coerce objects into type "numeric". Logical values `FALSE` and `TRUE` are turned into `0` and `1`, respectively. Characters are all turned to `NA`. Factor objects are coerced to numbers indicating their different levels. Results are not shown.
```{r, eval = FALSE}
as.numeric(logical_vec)
as.numeric(char_vec)
as.numeric(factor_vec)
```

Next, try to convert the character variable from character to factor to numeric; similarly, convert the factor variable from factor to character to numeric. As a result, the character variable can be converted to factor to numeric, and different characters are designated different numbers. Similarly, the factor variable can also be converted to character to numeric. 
```{r}
as.numeric(as.factor(char_vec))
as.character(factor_vec)
as.numeric(as.character(factor_vec))
```
However, in another case, if the factor vector is labeled with characters, it will first be converted to the label name, and therefore cannot be converted to numerics, so `NA` will be returned. See below: 
```{r}
factor_vec_labeled = factor(rep(c(0, 1), 5), levels = c(0, 1), labels = c("male","female"))
as.character(factor_vec_labeled)
as.numeric(as.character(factor_vec_labeled))
```


# Problem 2



